Motion planning for assessing accessibility of protein tunnels

### 2017-09-10 paper.tex
discussion

\section{Discussion}

The experiments show the advantage of the proposed motion planner. 
The planner outperforms the Retraction-based planner not only in the `easy' problems (with $\smin=0.5$), but also
in the case of difficult scenarios with larger ligands ($\smin=0.8$).

It is worth to mention, that the performance of both planners is influenced by their parameters, namely by
$\Imax$.
By increasing this parameter, we can expect better performance in finding the trajectories and consequently higher 
traversability rates, but with the increased computational time.

The internal performance of the planners is however not the main criteria.
Fom bio-chemical point of view, the main contribution of using motion planning for detecting the ligand trajectories is the ability to
consider

The proposed planner is very simple and it can be improved in many ways: by using a less aggressive scaling-down technique (in order
to obtain better trajectories) or by using only a subset of conformations and a better expansion (in order to speed it up).
Having a better motion planner (in the term of the speed or even success rate) however does not automatically mean that the resulting trajectories will help the biochemists more.
The presented work shows possible advantages of the analysis based on ligand trajectories.
By comparing the planned trajectories with trajectories observed in MD simulations, thresholds for the success rate or minimal tunnel throughput can be set, which will help biochemists to make decisions.

>>>>

\section{Discussion and utilization of the results}

The work in the paper is motivated by real needs of bio-chemists who need to analyze traversability of selected ligands in the tunnels.
To enable finding the trajectories, it is necessary to shrink the atomic radii of the ligands which has been utilized also in other tools~\cite{cortes2005path,guieysse2008structure}.
Finding trajectories for scaled-down ligands on one hand helps to cope with the narrow passages, but the question is, how realistic
such trajectories are.

Judging the accessibility of the active sites by the ligands based only on the traversability rates may be problematic if the ligands are scaled too much.
Instead of using a single version of the scale-down ligand, we only limit the minimal scale used by the planner (the parameter $\smin$).
The expansion step always prefers to expand the tree with the maximal scale, that leads to a new collision-free configuration.
The information about the scales used at the trajectory points (which can be extracted from the configuration tree of the proposed planner) brings the chemists new knowledge about the tunnels.
It allows them to detect parts of the tunnels, where the ligand has to scale-down significantly, and parts, where the ligand can be again
larger (Fig.~\ref{fig::dcp}).
According to the results of the motion planinng, DCP needs to shrink down to ca. 75~\% of its size (Fig.~\ref{fig::dcp}, point B).
However, as both the protein and the ligand adapt to each other, it is possible, that even a narrow part of the tunnel is traversable.
This was observed in the MD simulation of the DCP ligand in the 1CQW protein, which revealed that DCP can pass the first tunnel.
